SUPPORTING INFORMATION \\ Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-09 SI.tex
Deleted most old figures and large excess files

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{SUPPLEMENTAL INFORMATION \\ Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy}

@@

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.  The document should begin with an
%% abstract, if appropriate.  If one is given and should not be, the
%% contents will be gobbled.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

@@

% Fig. S1
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=0.9\textwidth, 
        scale=0.5]{Figs/figs_imp/allseq_crossval.pdf}
        \caption{Five-fold cross-validation of the SRV VAMP-2 scores to select the optimal number of SRV coordinates for each sequence. A knee in the VAMP-2 plot at approximately the fifth slow mode for each of the four sequences motivates the choice of a 5D projection. The absence of any significant separation in the training and testing VAMP-2 scores demonstrates that the 5D model is not overfitted.}
        \label{fig:allseq_srv_crossval}
	\end{center}
\end{figure}

@@

% Fig. S2
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=0.7\textwidth, 
        scale=0.5]{Figs/figs_imp/allseq_lags.pdf}
        \caption{Convergence of the MSM implied time scales $t_i$ as a function of lag time $\tau$. Solid lines indicate maximum likelihood result while dashed lines show the Bayesian ensemble means. The implied time scales for all sequences converge at a lag time of $\tau$ = 1.2 ns (vertical line). The black solid curve marks equality of the implied time scale and lag time and delimits the shaded region wherein the implied time scales are shorter than the lag time and cannot be resolved. \noter{(i) Eliminate tICA data. (ii) Eliminate legend. (iii) Add Bayesian ensemble uncertainties as colored shading behind the solid line. PyEMMA will do this calculation for you.}}
        \label{fig:all_seq_lags}
	\end{center}
\end{figure}

@@

% Fig. S3
\begin{figure}[ht!]
	\begin{center}
        %\includegraphics[width=0.7\textwidth, scale=0.5]{Figs/figs_imp/allseq_lags.pdf}
        \caption{Chapman-Kolmogorov (CK) tests comparing the probabilities of remaining within each macrostates for each sequence as a function of lag time predicted by an MSM constructed at the $\tau$ = 1.2 ns lag time (dashed blue line) versus those computed from an MSM constructed at the particular lag time (solid black line). The good agreement between these two results provides numerical validation of the Markovian nature of the $\tau$ = 1.2 ns lag time MSM.}
        \label{fig:all_seq_lags}
	\end{center}
\end{figure}

@@

% Fig. S4
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Figs/figs_imp/Tjump_fast_slow_fits.pdf}
        \caption{\noter{Can this be (i) simplified -- just show the left panels, the right panels are in the main text, (ii) report the kd values in the legends for the fits and provide the equation in the caption, and (iii) show data for all four seqs. Rewrite the caption and axes labels to reflect the terminology used in the main text (e.g., $k_d^\mathrm{slow}$, report Tm by writing it as text within white space of each figure for each sequence).}}
        \label{fig:fast_slow_fits}
	\end{center}
\end{figure}

@@

% Fig. S5
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=0.92\textwidth]{Figs/SX-1.pdf}
	\end{center}
\end{figure}
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=0.75\textwidth]{Figs/SX-2.pdf}
	\end{center}
\end{figure}
\begin{figure}[ht!]
	\begin{center}
        \caption{\noter{Add caption mirroring that from the main text.}}
        \label{fig:allseq_table}
	\end{center}
\end{figure}

@@

\rood{TBA -- full accounting complete with mathematics and references for model and its parameters of the application of NN model to predict free energy of each macrostate for each sequence} \cite{SantaLucia1998AThermodynamics}

@@

%We applied a similar diffusion map approach to investigate the structural composition of the F4 macrostate. We build diffusion maps using 10000 frames sampled from the macrostate (Figure \ref{fig:GC-core_dmaps}). Again, we were able to identify a combination of physical coordinates that closely correlated to the first two non-trivial diffusion modes. We found that a larger distance between internal A:T pairs increased the PCCA+ probability of inclusion into macrostate. This distance also correlated closely with the second diffusion map mode. Interestingly, we observed that the first non-trivial mode -- the feature that describes the most structural diversity in the system -- corresponds to difference in ``overlap'' distance between adjacent A or T bases and the GC core. We defined overlap as the distance between a G/C and the A/T adjacent to its native complement. At the extremes of this coordinate, one strand maintains most of its helical character while the other twists out of place. Because WC bonds are obstructed by the oligomer, the other strand is likely to fray before contacts can reform, or, less frequently, the duplex undergoes full dissociation. Interchange between these substates illustrate how the the F4 can maintain a longer lifetime compared to other A:T terminal sequences.

%\subsection{Explicit model comparisons}

@@

\bibliography{references}
%\bibliography{refs_mendeley2}

@@

\clearpage

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{SUPPORTING INFORMATION \\ Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy}

@@

\setlength{\marginparwidth}{2cm}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.  The document should begin with an
%% abstract, if appropriate.  If one is given and should not be, the
%% contents will be gobbled.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\newpage

@@

% Fig. S1
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=0.9\textwidth, 
        scale=0.5]{SIFig1.pdf}
        \caption{Five-fold cross-validation of the SRV VAMP-2 scores to select the optimal number of SRV coordinates for each sequence. A knee in the VAMP-2 plot was identified at approximately the fifth, fourth, third, and second slow modes for AT-all, GC-end, GC-core, and GC-mix, respectively. An embedding of corresponding dimensionality was then used to cluster frames into discrete states. The absence of any significant separation in the training and testing VAMP-2 scores demonstrates that model is not overfitted.}
        \label{fig:SIFig1}
	\end{center}
\end{figure}

@@

% Fig. S2
\begin{figure}[ht!]
	\centering
    \includegraphics[width=0.7\textwidth, scale=0.5]{SIFig2.pdf}
    \caption{Convergence of the MSM implied time scales $t_i$ as a function of lag time $\tau$. Solid lines indicate maximum likelihood result while dashed lines show the Bayesian ensemble means. The implied time scales for all sequences converge at a lag time of $\tau$ = 1.2 ns (vertical line). The black solid curve marks equality of the implied time scale and lag time and delimits the shaded region wherein the implied time scales are shorter than the lag time and cannot be resolved.}
    \label{fig:SIFig2}
\end{figure}

@@

% Fig. S3
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.1\textwidth]{SIFig3.pdf}
        \caption{Chapman-Kolmogorov (CK) tests comparing the probabilities of remaining within each macrostates for each sequence as a function of lag time predicted by an MSM constructed at the $\tau$ = 1.2 ns lag time (dashed blue line) versus those computed from an MSM constructed at the particular lag time (solid black line). The good agreement between these two results provides numerical validation of the Markovian nature of the $\tau$ = 1.2 ns lag time MSM.}
        \label{fig:SIFig3}
	\end{center}
\end{figure}

@@

% Fig. S4
\begin{figure}[ht!]
	\centering
    \includegraphics[width=.8\textwidth]{SIFig4.pdf}
    \caption{Exponential fits for both slow (dissociation) and fast (fraying) response for all four sequences during ``computational T-jump'' experiments. From 120 independent 1 $\mu$s simulations, we compiled the distribution of times at which terminal A:T and central base pairs first separate to a distance of of 1.3 nm. We extracted our computational estimate of $k_d^\mathrm{fast}$ by fitting a decaying exponential to the fraction of bound A:T termini as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$. Similarly, we extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$.}
    \label{fig:SIFig4}
\end{figure}

@@

% Fig. S5
\begin{figure}[ht!]
	\centering
    \includegraphics[width=0.8\textwidth]{SIFig5a.pdf}
\end{figure}
\begin{figure}[ht!]
	\centering
    \includegraphics[width=0.8\textwidth]{SIFig5b.pdf}
\end{figure}

\begin{figure}[ht!]
	\centering
    \caption{GC-mix hybridizes by nucleation-zippering and dehybridizes by fraying-peeling. Tracking of the 10 intermolecular distances between native WC base pairs over the course of four additional hybridization and dissociation events. Symmetrically permutable distances (e.g., A$_1$:T$_{10}$ and T$_{10}$:A$_1$) are reflected across the x-axis to avoid congestion in the plot. Circles superposed on the x-axis indicate the instantaneous MSM state assignment as dissociated D (blue) or hybridized H (orange). Hybridization tends to occur by a nucleation-zippering mechanism, wherein a native G:C WC pair and adjacent A:T pair or central A:T pairs first form prior to rapid formation of the complete duplex. Dehybridization tends to occur by a fraying-peeling mechanism wherein fraying of the two-base AT-tails on one or both sides of the duplex precedes dissociation of the central native base pairs and complete dissolution of the duplex.}
    \label{fig:SIFig5}
\end{figure}

@@

DNA duplex hybridization thermodynamics are most commonly predicted using nearest-neighbor (NN) models where the stability of a given base pair is assumed to solely depend on the identity and orientation of its adjacent base pairs. We apply the NN model reported by SantaLucia\citep{SantaLucia1998AThermodynamics}  to estimate the stability of fully intact, shifted, and frayed duplex configurations observed in our Markov State Model.

The NN model from SantaLucia assumes hybridization proceeds in a two-state all-or-nothing manner. The Gibbs free energy difference between the hybridized and dissociated states can be determined from calculation of NN enthalpy and entropy contributions.
\begin{equation}\label{SIe1}
\Delta G_{NN}^{\circ}(T) = \Delta H_{NN}^{\circ} - T\Delta S_{NN}^{\circ}
\end{equation}
Here,  and  are assumed to be independent of temperature.   and   are determined as the sum of all NN terms in a given configuration. 
\begin{equation}\label{SIe2}
\Delta H_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta H_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta H_{DE, j}^{\circ} + \Delta H_{init}^{\circ}
\end{equation}

\begin{equation}\label{SIe3}
\Delta S_{NN}^{\circ} = \sum_{i}^{n_{NN}}\Delta S_{NN, i}^{\circ} + \sum_{j}^{n_{DE}}\Delta S_{DE, j}^{\circ} + \Delta S_{init}^{\circ} + \Delta S_{sym}^{\circ}
\end{equation}

$\Delta H_{NN, i}^{\circ}$ is the NN enthalpy for a given dinucleotide step within duplex DNA. $\Delta H_{DE, j}^{\circ}$  corresponds to the enthalpic contribution from a given dangling end  base next to the duplex\citep{Santalucia2004TM}. $\Delta H_{init}^{\circ}$ is a duplex initiation term that accounts for terminal effects on duplex enthalpy. $\Delta S_{NN}^{\circ}$  is constructed from analogous NN, dangling end, and initiation terms in addition to a symmetry term, $\Delta S_{sym}^{\circ}$, which applies an entropic penalty to self-complementary sequences for maintaining C2 symmetry. The SantaLucia NN parameters are defined for a solution condition at neutral pH with a sodium concentration of 1M. We use the empirical salt correction developed by Owczarzy\citep{Owczarzy2008PredictingCations} to determine the NN parameters at the sodium concentration used for simulations in this work. The Owczarzy model assumes the affect of counterions on DNA hybridization thermodynamics is purely entropic.

\begin{equation}\label{SIe3}
\Delta S_{NN}^{\circ}(c_{Na^+}) = S_{NN}^{\circ}(1M) + \Delta H_{NN}^{\circ}[(4.29f_{GC}-3.95)\times 10^{-5} \ln{(c_{Na^+})}+ 9.4\times 10^{-6} (\ln{(c_{Na^+})})^2]
\end{equation}

$f_{GC}$ is the fractional G:C content of the duplex and $c_{Na^+}$ is the concentration of sodium counterions in M.

\begin{figure}[ht!]
	\centering
    \includegraphics[width=.85\textwidth]{SIFig6.pdf}
    \caption{Schematic of nearest neighbor contributions for an intact hybridized and out-of-register duplex with dangling ends (DE) stabilization.}
    \label{fig:SIFig4}
\end{figure}

@@

%We applied a similar diffusion map approach to investigate the structural composition of the F4 macrostate. We build diffusion maps using 10000 frames sampled from the macrostate (Figure \ref{fig:GC-core_dmaps}). Again, we were able to identify a combination of physical coordinates that closely correlated to the first two non-trivial diffusion modes. We found that a larger distance between internal A:T pairs increased the PCCA+ probability of inclusion into macrostate. This distance also correlated closely with the second diffusion map mode. Interestingly, we observed that the first non-trivial mode -- the feature that describes the most structural diversity in the system -- corresponds to difference in ``overlap'' distance between adjacent A or T bases and the GC core. We defined overlap as the distance between a G/C and the A/T adjacent to its native complement. At the extremes of this coordinate, one strand maintains most of its helical character while the other twists out of place. Because WC bonds are obstructed by the oligomer, the other strand is likely to fray before contacts can reform, or, less frequently, the duplex undergoes full dissociation. Interchange between these substates illustrate how the the F4 can maintain a longer lifetime compared to other A:T terminal sequences. 
%\subsection{Explicit model comparisons}

@@

\bibliography{references_210505}
%\bibliography{refs_mendeley2}

@@

%\clearpage
